---
title: "HW3 4355"
author: "Courtney Schaller"
date: "2/25/2021"
output: pdf_document
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
knitr::opts_chunk$set(echo = TRUE)
```

3. (R programming) The weight and systolic blood pressure of 26 randomly selected males in the age group 25–30 are found in the attached CSV file. 
We fit a simple linear regression model relating systolic blood pressure (y)
to weight (x). 

(a) Draw the scatterplot of the data points with R function, plot. Add the linear regression line using the R
function abline after plot. (10 points)
Note: Present two plots: one is a scatterplot only and the other is a scatterplot with the regression line.

```{r}
g <- "C:\\Users\\Schaller\\Downloads\\hw2_systolic_bp.csv"
df <- read.csv(g, header = TRUE)
plot(df$ï..weight, df$sys.bp)
plot(df$ï..weight, df$sys.bp)
abline(lm(formula = df$sys.bp ~ df$ï..weight))
```

(b) Find the coefficient of determination (R2) and compute the total sum of squares (SST). Using the two, compute the regression sum of squares (SSR). (15 points)

```{r}
#R2 * SST = SSR
model <- lm(sys.bp ~ ï..weight, data = df)
R2 <- summary(model)$r.squared
SST <- sum((df$sys.bp - mean(df$sys.bp))^2)
SSR <- R2 * SST

R2
SST
SSR
```

(c) Find the 99 % confidence interval on (i) the slope and the intercept, (ii) the mean response at x = 160, 180, 200, and (iii) 26 future observations whose weights are identical to the males’ in the dataset. (15 points) Hint: Create a data.frame, where x values are 160, 180, 200 for (ii) and all the 26 males’ weights for (iii). Specify an argument of confint or predict, level, to be 0.99. 

```{r}
confint(model, level=0.99)
tests <- data.frame(ï..weight = c(160, 180, 200))

predictions <- data.frame(predict(model, df, interval = 'prediction', level = 0.99))
testpredictions <- data.frame(predict(model, tests, interval = 'confidence', level = 0.99))

cbind(tests, testpredictions)
pred.bp <- cbind(df, predictions)
pred.bp
```

(d) Merge the data.frame object of the 26 data points with the prediction outcome object from (c) using R function cbind. Name the data.frame object of size 26 × 5 created pred.bp. Note: Present two plots: one is a scatterplot with the interval curve and the other is a scatterplot with
both the interval and prediction curve. (10 points)

```{r}
pred.bp <- cbind(df, predictions)
pred.bp
#Draw the confidence interval curve using the commands below.
library("ggplot2")
p <- ggplot(pred.bp, aes(ï..weight, sys.bp)) +
geom_point() +
stat_smooth(method = lm)
p
#Draw the prediction interval curve using the commands below.
p + geom_line(aes(y = lwr), color = "red", linetype = "dashed")+
geom_line(aes(y = upr), color = "red", linetype = "dashed")
```

(e) Attach the R code and console output you used for (a)-(d). (5 points)
