---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---
```{r}
library(tidyverse)
library(ggplot2)
```

1.$$y=\beta_{0}+\beta_{1} x_{1}+\beta_{2} x_{2}+\beta_{3} x_{3}+\varepsilon$$
(a) $$H_{0}: \beta_{1}+2 \beta_{2}=3$$
$$D=\begin{bmatrix}
 0 & 1 & 2 & 0
\end{bmatrix}$$
$$d = 3$$
(b) $$H_{0}: \frac{\beta_{1}+\beta_{2}}{2}=\beta_{3}$$
$$D=\begin{bmatrix}
 0 & 1 & 1 & -2
\end{bmatrix}$$
$$d = 0$$
(c) $H_{0}: \beta_{1}=\beta_{2}, \beta_{3}=\beta_{4}$
$$D=\begin{bmatrix}
 0 & 1 & 0 & -1 & 0\\
 0 & 0 & 1 &  0 & -1
\end{bmatrix}$$
$$d = \begin{bmatrix}
 0 \\
 0
\end{bmatrix}$$

(d) $$H_{0}: \beta_{1}-2 \beta_{2}=4 \beta_{3}, \beta_{1}+2 \beta_{2}=0$$

$$\begin{bmatrix}
 0 & 1 & -2 & -4\\
 0 & 1 & 2 &  0
\end{bmatrix}$$

$$d = \begin{bmatrix}
 0 \\
 0
\end{bmatrix}$$

(e) $H_{0}: \beta_{1}=\beta_{2}=\beta_{3}=\beta_{4}$

$$\begin{bmatrix}
 0 & 1 & -1 &  0 & 0\\
 0 & 0 &  1 & -1 & 0\\
 0 & 0 &  0 &  1 & -1
\end{bmatrix}$$

$$d = \begin{bmatrix}
 0 \\
 0\\
 0
\end{bmatrix}$$
a.
```{r}
football <- read.csv(
  "https://raw.githubusercontent.com/schourtney/stat4355/main/football.csv",
                     header=T)

x <- as.vector(colnames(football))[-c(1, 2)]
ggplot(football, aes(x = .data[[x[1]]], y)) + geom_point()
ggplot(football, aes(x = .data[[x[2]]], y)) + geom_point()
ggplot(football, aes(x = .data[[x[3]]], y)) + geom_point()
ggplot(football, aes(x = .data[[x[4]]], y)) + geom_point()
ggplot(football, aes(x = .data[[x[5]]], y)) + geom_point()
ggplot(football, aes(x = .data[[x[6]]], y)) + geom_point()
ggplot(football, aes(x = .data[[x[7]]], y)) + geom_point()
ggplot(football, aes(x = .data[[x[8]]], y)) + geom_point()
ggplot(football, aes(x = .data[[x[9]]], y)) + geom_point()
```

b.
```{r}
lm <- lm(y ~ x1 + x2 + x5 + x7 + x8, data = football)
summary(lm)
variance <- (summary(lm)$sigma)^2
```
The fitted linear model is $Y = -0.5165 + 0.0008X_1 + 0.0035X_2 + 0.0104X_5 + 0.1337X_7 - 0.0037X_8$
Variance is $1.768^2$, or 3.1270
$R^2 = 0.7896$, adjusted $R^2 = 0.7418$

c.
The parameter estimates $\hat{\beta_{k}}$ are such that with each unit increase in $X_1, X_2, X_5, \dots$, we expect an increase to y, number of games won, by $\hat{\beta_1}, \hat{\beta_2}, \hat{\beta_3}, \dots$, respectively. e.g. one increased rushing yard ($X_1$) would increase games won by $\hat{\beta_1}$. $\hat{\sigma}^2 = 3.1270$ is the variance in our predicted number of games won.

d.
```{r}
anova(lm)
SSreg <- sum(115.068, 83.343, 3.228, 18.051, 38.381)
MSreg <- SSreg/5
SourceOfVariation <- c("Regression", "Residual", "Total")
SS <- c(SSreg, 68.794, SSreg+68.794)
DF <- c(5, 22, 22+5)
MS <- c(MSreg, 3.127, "")
F <- c(MSreg/3.127, "", "")
df <- data.frame(SourceOfVariation, SS, DF, MS, F)
df
```
e.
```{r}
fit0 <- lm(y~1, data = football)
anova(fit0, lm)
SourceOfVariation <- c("Regression", "Residual", "Total")
SS <- c(326.96-68.79, 68.79, 326.96)
DF <- c(5, 22, 27)
MS <- c(258.17/5, "", "")
F <- c(16.512, "", "")
df <- data.frame(SourceOfVariation, SS, DF, MS, F)
df
```
f.
Yes, the two tables are the same. Our 5-predictor model has a p-value of 8.363e-07<0.05, so we reject the null hypothesis and conclude that there is a linear relationship between the 5 X inputs (rushing yards, passing yards, etc.) and the number of games won (Y). Among these inputs, all appear to have a significant effect on Y except X1, X5, and X7 with p-values >0.05 (0.623008, 0.815766, and 0.341628, respectively.)

g.
(i). null hypothesis: $\hat{\beta_1}=0$
alternative hypothesis: $\hat{\beta_1}\neq0$ \newline
(ii). T-distribution, we will reject the null if our calculated t value is outside the bounds of $t_{0.025, 22}$ \newline
(iii). t = 0.499, p = 0.623008, we fail to reject the null hypothesis and conclude the effect of passing yardage on overall games won is insignificant. There doesn't appear to be a linear relationship between these two variables.

h.
```{r}
lm2 <- lm(y ~ x2 + x8, data = football)
summary(lm2)
```
(i). null hypothesis: $\hat{\beta_2} = \hat{\beta_8} = 0$
alt hypothesis: $\hat{\beta_2}, \hat{\beta_8} \neq 0$ \newline
(ii). F-distribution, we will reject the null if our calculated f-statitstic is outside the bounds of $f_{\alpha = 0.05, f1=2, f2=25}$ \newline
(iii). f-statistic is 36.19, p-val is 4.152e-08<0.05, we reject the null hypothesis and determine that there is a linear relationship between X2,X8 and Y.

i. While R^2 is higher for the 5-predictor model, and therefore the model /could/ be better at predicting true number of games won, the interpretability of the 2-predictor model is better, and as the effect of the three extra predictors isn't significant, I prefer the 2-predictor model to avoid overfitting to training data.

